Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-13 paper/paper.tex
dots

\mathchardef\mhyphen="2D
\newcommand{\OPT}{\mathit{OPT}}
\newcommand{\ALG}{ALG}
\newcommand{\OBRP}{OBRP}
\newcommand{\PPOBRP}{PPOBRP}

@@

\author{Maciej Pacut}
%\orcid{0000-0002-6379-1490}
\affiliation{%
  \department{Faculty of Computer Science}
  \institution{University of Vienna}
  \country{Austria}
}

\author{Mahmoud Parham} 
\affiliation{%
  \department{Faculty of Computer Science}
  \institution{University of Vienna}
  \country{Austria}
}

\author{Stefan Schmid} 
\affiliation{%
  \department{Faculty of Computer Science}
  \institution{University of Vienna}
  \country{Austria}
}

@@

Additionally, we contribute to the research on the optimal online algorithm by presenting a $O(k\ell)$-competitive algorithm for the special case
where the communication graph (i.e., request sequence) admits a \emph{perfect partition},
i.e., a (multi-way) cut of cost zero that is also an optimal partition for the offline problem.
%and every algorithm must colocate every communicating pair.
Note that our $\Omega(k\ell)$ lower bound generates the input sequence
so that the optimal offline choice is a perfect partition.
\maciek{Also, it is easy to construct the adversary that forces collocation of requested nodes}
\mahmoud{this is already the case in the lower bound proof. Do we still need the last sentence here?}
\maciek{In general it is OK to emphasize some parts of the paper in the intro. Not sure if this sentence in particular is that important, though.}

@@

We assume that the communication patterns are not known in advance to our algorithm.
We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.

We assume that the physical substrate network is a~tree of height one.
That is, every physical machine (leaf) is connected directly to the root (that has no hosting capabilities).
A single physical machine hosts a fixed number of virtual machines.
The model restricted to such networks becomes a variant of online graph clustering.
That is, we are given a~set of~$n$ nodes (virtual machines) with time-varying pairwise
communication patterns, which have to be partitioned into~$\ell$~physical machines, each of
capacity $k=n/\ell$.

@@

\maciek{I suggest to rename the problem to either: (1) Dynamic Balanced Graph Partition or (2) Online Balanced Graph Partition, as in the SIDMA paper}

Formally, the online \emph{Balanced Re-Partitioning} problem (\OBRP{}) is defined as
follows. We are given a set of $n$ nodes,
initially arbitrarily partitioned into $\ell$~clusters,
each of size~$k$.
We say that nodes in $C \subset V$ are \emph{collocated} in a partition $P$
if $P$ assigns them to the same cluster.
\maciek{We can ommit the formal definition: "I.e.,
$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". However, if we use the notion of cluster later, then TODO: define cluster}
The input consists of a sequence of pairwise communication requests
$\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a fixed amount of data.
Before serving the request,
the online algorithm
can re-partition,
i.e.,
it may (optionally) move some of the nodes into clusters different than their current assigned clusters.
Once the re-partition is executed,
we serve
a~communication request between two  nodes at cost~0 if they are collocated,
or at cost~1 if they are located in different clusters.
The cost of migrating a node from one cluster to another
is~$\alpha \in \mathbb{Z}^+$.
For any algorithm $ALG$,
its cost,
denoted by $ALG(\sigma)$,
is the total cost of communications and
the cost of migrations performed by $ALG$ while serving the sequence $\sigma$.
\maciek{Not an expert on commas in english, but here it feels like we have too much. Use some tool to determine (Grammarly?)}

@@

Given this sequence of requests,
the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
 swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
Hence,
OPT pays $4$ for node migrations and
ALG pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k.\ell$ nodes,
therefore $|X| \geq k.\ell - k - \ell \in \Omega(k\cdot\ell)$.
\end{proof}

@@

The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\ell^2)$.

@@

\todo{Make ORCID visible in ACM style}

>>>>

\mathchardef\mhyphen="2D
\newcommand{\OPT}{\mathit{OPT}}
\newcommand{\ALG}{ALG}
\newcommand{\OBRP}{BRP}
\newcommand{\PPOBRP}{PP-BRP}

@@

\author{Maciej Pacut}
\orcid{0000-0002-6379-1490}
\affiliation{%
  \institution{Faculty of Computer Science, University of Vienna}
  \country{Austria}
}

\author{Mahmoud Parham} 
\affiliation{%
  \institution{Faculty of Computer Science, University of Vienna}
  \country{Austria}
}

\author{Stefan Schmid} 
\affiliation{%
  \institution{Faculty of Computer Science, University of Vienna}
  \country{Austria}
}

@@

Additionally, we contribute to the research on the optimal online algorithm by presenting a $O(k\cdot \ell)$-competitive algorithm for the special case
where the communication graph (i.e., request sequence) admits a \emph{perfect partition},
i.e., a (multi-way) cut of cost zero that is also an optimal partition for the offline problem.
%and every algorithm must colocate every communicating pair.
Note that our $\Omega(k\cdot \ell)$ lower bound generates the input sequence
so that the optimal offline choice is a perfect partition.
%\maciek{Also, it is easy to construct the adversary that forces collocation of requested nodes}
%\mahmoud{this is already the case in the lower bound proof. Do we still need the last sentence here?}
%\maciek{In general it is OK to emphasize some parts of the paper in the intro. %Not sure if this sentence in particular is that important, though.}

@@

We assume that the communication patterns are not known in advance to our algorithm.
We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
t
We assume that the physical substrate network is a~tree of height one.
That is, every physical machine (leaf) is connected directly to the root (that has no hosting capabilities).
A single physical machine hosts a fixed number of virtual machines.
The model restricted to such networks becomes a variant of online graph clustering.
That is, we are given a~set of~$n$ nodes (virtual machines) with time-varying pairwise
communication patterns, which have to be partitioned into~$\ell$~physical machines, each of
capacity $k=n/\ell$.

@@

Formally, the online \emph{Balanced Graph RePartitioning} problem (\OBRP{}) is defined as
follows. We are given a set of $n$ nodes,
initially arbitrarily partitioned into $\ell$~clusters,
each of size~$k$.
We say that nodes in $C \subset V$ are \emph{collocated} in a partition $P$
if $P$ assigns them to the same cluster.
\maciek{We can ommit the formal definition: "I.e.,
$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". However, if we use the notion of cluster later, then TODO: define cluster}
The input consists of a sequence of pairwise communication requests
$\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a fixed amount of data.
Before serving the request,
the online algorithm
can re-partition,
i.e.,
it may (optionally) move some of the nodes into clusters different than their current assigned clusters.
Once the re-partition is executed,
we serve
a~communication request between two  nodes at cost~0 if they are collocated,
or at cost~1 if they are located in different clusters.
The cost of migrating a node from one cluster to another
is~$\alpha \in \mathbb{Z}^+$.
For any algorithm $ALG$,
its cost,
denoted by $ALG(\sigma)$,
is the total cost of communications and
the cost of migrations performed by $ALG$ while serving the sequence $\sigma$.
\maciek{Not an expert on commas in english, but here it feels like we have too much. Use some tool to determine (Grammarly?)}

@@

Given this sequence of requests,
the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
 swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
Hence,
OPT pays $4$ for node migrations and
ALG pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
therefore $|X| \geq k \cdot \ell - k - \ell \in \Omega(k\cdot\ell)$.
\end{proof}

@@

The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\cdot\ell^2)$.

@@

\todo{Make ORCID visible in ACM style: "Right now we do not typeset ORCIDs" (Source: ACM Style Docs)}
